```{r configuration, echo=FALSE,message=FALSE,results='hide'}
#
# Global settings
#
url <- getOption("opal.url", "https://demo.obiba.org:8443")
username <- getOption("opal.username", "administrator")
password <- getOption("opal.password", "password")
datasource <- getOption("opal.datasource", "mica_demo")
table <- getOption("opal.table", "FNAC")
withStatistics <- getOption("opal.withStatistics", TRUE)
withGraphs <- getOption("opal.withGraphs", TRUE)
```

```{r set-up, echo=FALSE,message=FALSE,results='hide'}
# set up dictionary and data
library(opal)
o <- opal.login(username, password, url)
variables <- opal.variables(o, datasource, table)
t <- opal.table(o, datasource, table, counts=TRUE)
if (withStatistics | withGraphs) {
  opal.execute(o, 'require(e1071)')
  opal.assign(o,'D', paste0(datasource, ".", table))
}
if (withGraphs) {
  require(ggplot2)
  opal.execute(o, 'require(ggplot2)')
}
```

# `r table` Data Dictionary

<br/>

Name of the report     | Data Dictionary for `r table`
---------------------- | -------------------
Project                | `r datasource`
Table                  | `r table`
Number of variables    | `r t$variableCount`
Number of participants | `r t$valueSetCount`
Date                   | `r as.character(Sys.Date())`

## Variables

```{r variables, results='asis', echo=FALSE}
for (variable in variables) {
  cat(paste0("* [", variable$name, "](#", gsub("\\.","_",variable$name),")\n"))
}
```

```{r run-variable-md, include=FALSE}
out <- NULL
for (variable in variables) {
  out <- c(out, knit_child('opal-variable.Rmd'), sep='\n')
}
```

`r paste(out, collapse='\n')`

```{r clean-up, results=FALSE,echo=FALSE}
opal.logout(o)
```

<style type="text/css">
@media print{
  body {
    font-size: 12px;
  }
  h2, h3 {
    page-break-before: always;
  }
}

@media screen{
  h3 {
    border-top: solid 1px #ccc;
    padding-top: 20px;
  }
}

.navbar {
  display: none;
}

#toc {
  overflow: overlay;
  max-height: 500px;
}

.table {
  width: 0%;
}

.table td {
  min-width: 200px;
}

</style>