---
title: "Opal Projects API"
author: "Yannick Marcon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Opal Projects API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[Opal](https://www.obiba.org/pages/products/opal/) stores data and meta-data in projects that are accessible through web services. See the [Variables and Data](http://opaldoc.obiba.org/en/latest/variables-data.html) documentation page that explains the data model of Opal. The Opal R package exposes projects related functions:

* list projects,
* list tables,
* list variables,
* list taxonomies,
* and functions for getting all associated details (of a project, a table, a variable etc.) ...

Note that these functions do not create a R session on the server side: it is only accessing the content of the Opal server (permission checks apply).

Setup the connection with Opal:

```{r eval=FALSE}
library(opalr)
o <- opal.login("administrator", "password", url = "https://opal-demo.obiba.org")
```

## Project Operations

List the projects:

```{r eval=FALSE}
opal.projects(o)
```

List the datasources (there is one per project) to get the details about how the data are persisted and what are the tables:

```{r eval=FALSE}
opal.datasources(o)
```

List the tables in a datasource, with their count of variables and entities:

```{r eval=FALSE}
opal.tables(o, datasource = "CPTP", counts = TRUE)
```

## Table Operations

The table object can be retrieved as follow:

```{r eval=FALSE}
opal.table(o, "CPTP", "Coreqx_final", counts = TRUE)
```



### Table Dictionary Operations

List the variables of a table and get the details of the variable annotations (one column per variable attribute with namespace). This list also includes the category properties:

```{r eval=FALSE}
opal.variables(o, "CPTP", "Coreqx_final")
```

It is also possible to get the full data dictionary of a table, as separate data frames of variables and categories. This is the recommended format for working with a data dictionary:

```{r eval=FALSE}
dico <- opal.table_dictionary_get(o, "CPTP", "Coreqx_final")
dico$variables
dico$categories
```

### Table Values Operations

Get the values in a table for a specific entity:

```{r eval=FALSE}
opal.valueset(o, "FNAC", "FNAC", identifier = "1000502517535681229")
```

It is possible to truncate a table, i.e. delete ALL the values of a table (which must not be a view):

```{r eval=FALSE}
opal.table_truncate(o, "FNAC", "FNAC")
```

## Taxonomy Operations

Variable attributes can be described by taxonomies. List the taxonomies:

```{r eval=FALSE}
opal.taxonomies(o)
```

List the vocabularies of a taxonomy:

```{r eval=FALSE}
opal.vocabularies(o, taxonomy = "Mlstr_area")
```

List the terms of a vocabulary:

```{r eval=FALSE}
opal.terms(o, taxonomy = "Mlstr_area", vocabulary = "Lifestyle_behaviours")
```



Good practice is to free server resources by sending a logout request:

```{r eval=FALSE}
opal.logout(o)
```

