---
title: "Opal Files API"
author: "Yannick Marcon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Opal has an internal file system that is accessible through web services. The opal R package exposes files management related functions:

* upload a file,
* download a file or folder (optionaly encrypted),
* move or rename a file or folder,
* create a folder,
* remove a file or folder,
* write a file to the remote R session workspace,
* read a file from the remote R session workspace.

Setup the connection with Opal:

```{r}
library(opalr)
o <- opal.login('administrator', 'password', 'https://opal-demo.obiba.org')
```

Download a file:

```{r}
opal.file_download(o, '/tmp/datashield.zip')
```

Download a file, protected by a password:

```{r}
opal.file_download(o, '/tmp/datashield.zip', 'datashield-encrypted.zip', key='ABCDEFGHIJKL')
```

Upload the file at another location:

```{r}
list.files()
opal.file_upload(o, 'datashield.zip', '/projects/datashield')
```

Create a folder and list folder content:

```{r}
opal.file_mkdir(o, '/projects/datashield/foo')
opal.file_ls(o, '/projects/datashield')
```

Move file to the new folder and list folder content:

```{r}
opal.file_mv(o, '/projects/datashield/datashield.zip', '/projects/datashield/foo')
opal.file_ls(o, '/projects/datashield/foo')
```

Rename the new folder and list folder content:

```{r}
opal.file_mv(o, '/projects/datashield/foo', '/projects/datashield/bar')
opal.file_ls(o, '/projects/datashield')
```

Write a file from the Opal file system into the R server session workspace:

```{r}
opal.file_write(o, '/projects/datashield/bar/datashield.zip')
opal.execute(o, 'list.files()')
```

Read back the file into the Opal file system:

```{r}
opal.file_read(o, 'datashield.zip', '/projects/datashield/bar/ds.zip')
opal.file_ls(o, '/projects/datashield/bar')
```


Remove created folder and list folder content:

```{r}
opal.file_rm(o, '/projects/datashield/bar')
opal.file_ls(o, '/projects/datashield')
```

Good practice is to free server resources by sending a logout request:

```{r}
opal.logout(o)
unlink('datashield-encrypted.zip')
unlink('datashield.zip')
```

