---
title: "Opal R Session API"
author: "Yannick Marcon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Opal R Session API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

[Opal](https://www.obiba.org/pages/products/opal/) is connected with a R server. Each Opal user having the appropriate permissions can create a R session on the server side through Opal. See the [R Server](http://opaldoc.obiba.org/en/latest/admin/rserver.html) documentation page that explains the relationship beteen Opal and the R server. The Opal R package exposes data management related functions:

* assignment of Opal data into the remote R session,
* execution of R expressions on server side,
* management of R workspace image.

Note that these functions do create a R session on the server side.

Setup the connection with Opal:

```{r eval=FALSE}
library(opalr)
o <- opal.login("administrator", "password", "https://opal-demo.obiba.org")
```

Assign a Opal table to a *data.frame* (if user has permission to see values of the Opal table):

```{r eval=FALSE}
opal.assign.table(o, symbol = "df", value = "datashield.CNSIM1")
```

In place of a standard *data.frame*, Opal can assign data into a [tibble](https://tibble.tidyverse.org/) which has the advantage of containing the data dictionary:

```{r eval=FALSE}
opal.assign.table.tibble(o, symbol = "tbl", value = "datashield.CNSIM1")
```

List the R symbols that lives in the remote R session:

```{r eval=FALSE}
opal.symbols(o)
```

R expressions can be executed on the server side:

```{r eval=FALSE}
opal.execute(o, script = "names(tbl)")
```

The remote data can be downloaded into the client side R session:

```{r eval=FALSE}
tbl <- opal.execute(o, script = "tbl")
```

The image of the remote R session can be saved in a *workspace* (to be restored at login time):

```{r eval=FALSE}
opal.workspace_save(o, save="cnsim1")
opal.workspaces(o)
```


Good practice is to free server resources by sending a logout request:

```{r eval=FALSE}
opal.logout(o)
```

